A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-13 sections/Overview.tex
polish language in 2.1 + comment

The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.

@@

In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
we give the subtyping and typing relations for the declarative system.

\paragraph{Overlapping rules}
\jimmy{Move to the Alg. Section}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \forall a. A \le \forall a. B$.
In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.

>>>>

The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
A declarative type $A$ is either the unit type $1$, a type variable $a$,
a universal quantification $\forall a. A$ or a function type $A \to B$.
Nested universal quantifiers are allowed in types,
but monotypes $\tau$ do not contain any universal quantifier.
Contexts $\Psi$ are just sequences of type variable declarations.

@@

Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
give the subtyping and typing relations for the declarative DK system.
The DK subtyping relation was adopted from \citet{odersky1996putting}.

\paragraph{Declarative Subtyping}
Figure~\ref{fig:decl:sub} shows DK's declarative subtyping judgment $\Psi \vdash A \le B$,
which was adopted from \citet{odersky1996putting}. It compares the
degree of polymorphism between between $A$ and $B$ in DK's implicit polymorphic type system. 
Essentially, if $A$ can always be instantiated to match any instantiation of $B$,
then A is ``at least as polymorphic as'' $B$. We also 
say that $A$ is ``more polymorphic than'' $B$ and write $A \le B$.

Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
handle simple cases that do not involve universal quantifiers.
The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
being covariant on the return type and contravariant on the argument type.
Rule $\mathtt{{\le}\forall R}$ states that if $A$ is a subtype of $B$
in the context $\Psi, a$, where $a$ is fresh in $A$, then $A\le\all B$.
Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
already indicates that $\all B$ is a general type),
then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.

The most interesting rule is $\mathtt{{\le}\forall L}$.
If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
then $\all A \le B$.
The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
declarative rule, but the algorithmic system does not guess and defers the
instantiation until it can determine the monotype deterministically.
The fact that $\tau$ is a monotype rules out the possibility of
polymorphic (or impredicative) instantiation.\tom{say why this restriction is imposed}
Peyton Jones et al.~\cite{} also impose the restriction of
predicative instantiation in their type system.
Both systems are adopt by several practical programming languages.

\paragraph{Declarative Typing}
\jimmy{TODO decl. typing description}

\paragraph{Overlapping Rules}
\jimmy{Move to the Alg. Section}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \all A \le \all B$.
In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.
